---
title: "UTH wifi usage analysis"
author: "Xuan Hieu Nguyen"
format:
  html:
    css: style.css
editor: visual
execute:
  warning: false
  message: false
code-fold: true
---

The codes to read in the data:

```{r}
# 1. Load the necessary libraries
library(readxl)
library(tidyverse)

# Event log
event_log <- read_csv("Event_log.csv") |>
  janitor::clean_names() |> # clean column names
  mutate(time_07 = ymd_hms(paste("2025", time_07))) # Turn `time_07` column into Date object

# 2. Set the path to your file
file_path <- "Summary Report 2025-01-01 - 2025-06-30.xlsx"

# 3. Get the names of all sheets in the file
sheet_names <- excel_sheets(file_path)

# 4. Read all sheets into a list of data frames
all_sheets <- set_names(sheet_names) %>% 
  map(~ read_excel(path = file_path, sheet = .x) |>
        dplyr::rename_with(tolower) |>
        janitor::clean_names())
```

# I. The event log (.csv file)

## Log time

First of all, I was curious about the time interval of `event_log`, so I had a look at it:

```{r}
paste("Log time starts:", min(event_log$time_07)) |> cat()
paste("Log time ends:", max(event_log$time_07)) |> cat()
max(event_log$time_07) - min(event_log$time_07)
```

So, `event_log` contains **1000 records** within **32 seconds**!!!

## Access point

```{r}
event_log <- event_log |>
  mutate(zone = str_sub(access_point, 1, 4), .after = access_point)

event_log$access_point |>
  table(useNA = "ifany") |>
  as.data.frame() |>
  arrange(-Freq) -> seeing_access_point

event_log$zone |>
  table(useNA = "ifany") |>
  as.data.frame() |>
  arrange(-Freq) -> seeing_zone
```

## SSIDs

```{r}
SSID_summary <- event_log$ssid |>
  table(useNA = "ifany") |>
  as.data.frame() |>
  rename("SSID" = Var1, "Count" = Freq)

knitr::kable(
  SSID_summary,
  caption = "All SSIDs and their counts from Event_log.csv"
)
```

Out of a total of **`r SSID_summary$Count |> sum()`** SSIDs, **`r SSID_summary |> filter(is.na(SSID)) |> pull(Count)`** of them are **NAs** (blank). I had a talk about this with **chatGPT** ([Link to the chat](https://chatgpt.com/share/694215c0-55d8-800a-8662-fde501d27b85)), it suggested that perhaps the blank SSIDs might indicate the use of **cellular networks** (5G, LTE, etc.).

## Clients

Clients can be categorized into different groups:

-   Clients with **device-like names** (e.g., `MacBook` or `Android-…`), will be grouped into their respective devices, so `MACs-MacBook-Pro-2` belongs to `Mac`, whereas `Android-5` belongs to `Android`.

-   **6 bytes, hex, colon-separated** clients, such as `a6:06:1f:a7:6f:70` and `76:95:75:d9:2d:32`, are **MAC addresses** and will be put into `Mac` group.

-   Clients with `DESKTOP` or `LAPTOP` leading in their names (for example, `DESKTOP-7U8VR5E` and `LAPTOP-2O7TSJGK`) will be infered as `Windows`.

-   Clients with **32-digit number strings** like `1050b2c6-1e4d-43a6-b92e-ebad95e513e6` and `61a5ff55-e983-4087-9e4e-f91935a9e335` are **UUIDs (universally unique identifiers)** and will be labelled as such (These clients can be classified into other groups, but there is no way to tell which devices they are).

-   Clients that do not belong to any groups above, like `Tui-la-Huyen-neee`, `Nguoi-dung-an-danh`, etc. will be labelled as `Other` (They may belong to one of the groups above, or an entirely different group).

The following plot shows the **unique count** of different client types (There can be multiple records of 1 client):

```{r}
device_names <- "android|iphone|ipad|mac|windows|oppo|poco|xiaomi|laptop|redmi|samsung|galaxy|realme|nokia"

event_log <- event_log |>
  dplyr::mutate(
    client_type = dplyr::case_when(
      grepl("^[0-9a-f]{2}(:[0-9a-f]{2}){5}$", client, ignore.case = TRUE) ~ "Mac",
      grepl("^[0-9a-f\\-]{36}$", client, ignore.case = TRUE) ~ "UUID",
      grepl("DESKTOP|LAPTOP", client, ignore.case = FALSE) ~ "Windows",
      str_detect(
        client, regex(device_names, ignore_case = TRUE)
      ) ~ str_extract(
        client, regex(device_names, ignore_case = TRUE)
      ) |> str_to_title(),
      TRUE ~ "Other"
    ), .after = client
  )

event_log |>
  select(client, client_type) |>
  distinct() |> # remove duplicated clients
  pull(client_type) |>
  table(useNA = "ifany") |>
  as.data.frame() |>
  arrange(-Freq) |>
  mutate(Var1 = fct_reorder(Var1, Freq, .desc = TRUE)) |>
  rename("Client type" = Var1, "Count" = Freq) -> unique_client_type

p_client_type <- ggplot(unique_client_type, aes(`Client type`, Count)) +
  geom_col(fill = "steelblue", color = "skyblue") +
  labs(title = "Unique client counts by client types") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

plotly::ggplotly(p_client_type)
```

Most unique clients were **Macs**, followed by **UUIDs** and **Others**. As mentioned, `UUID` and `Other` can be any other groups, so **Macs** may not truly be the majority count.

## Category, event type & detials

### Category & event type

The `category` and `event_type` columns are closely related. Here you can see there are only **7 unique combinations** when I picked out these 2 columns and remove duplicated combinations:

```{r}
event_log |>
  select(category, event_type) |> # select only `category` and `event_type` columns
  distinct() |> # remove duplicated pairs
  arrange(event_type) |>
  knitr::kable(caption = "All unique category + event_type combinations from Event_log.csv")
```

```{r}
event_log |>
  select(event_type, details) |>
  group_by(event_type) |>
  summarise(top_10_unique_details = details |> unique() |> head(10)) |>
  knitr::kable() |>
  kableExtra::collapse_rows()
```

### Details: Events dropped

I want to have a look at **"events dropped"** category specifically. Each "events dropped" has a specific count, for example, **"287 events were not logged. <...>"**. I will extract those numbers and sum them up by **zones** and **access points**:

```{r}
events_dropped <- event_log |>
  filter(event_type == "Events dropped") |>
  mutate(drop_count = str_extract(details, "\\d+") |> as.integer())

dropped_sum <- events_dropped |>
  group_by(zone, access_point) |>
  summarise(count = sum(drop_count))

p_jitter <- ggplot(dropped_sum, aes(zone, count, colour = access_point)) +
  geom_jitter(width = 0.2) +
  labs(title = "Jitterplot of dropped event counts") +
  theme_bw()

plotly::ggplotly(p_jitter)

dropped_sum_zone <- events_dropped |>
  group_by(zone) |>
  summarise(total = sum(drop_count))

knitr::kable(dropped_sum_zone)
```

# II. The Summary Report (.xlsx file)

```{r}
get_client <- function(x) {
  event_log |>
    filter(client == x) |>
    arrange(time_07) |>
    list()
}

individual_client_log <- event_log |>
  group_by(client) |>
  summarise(client_log = get_client(client),
            row_count = map_int(client_log, nrow))

head(individual_client_log)
```

Cross-matching clients in CSV and XLSX files:

```{r}
unique_clients <- event_log |>
  select(client, client_type) |>
  distinct()

client_matching <- inner_join(
  unique_clients, all_sheets$`Top clients by usage`,
  by = join_by("client" == "description")
)
```
